---
title: "Assignment 6"
author: "Inaya Ahmed"
format: html
editor: visual
---

## DATA-613 Assignment 6

### Part I: Answer the following question by running SQL using RStudio. 

```{r}
# Loading libraries
library(DBI)
library(duckdb)
library(nycflights13)

# Downloading duckdb file
download.file(
  "https://data-science-master.github.io/lectures/data/flights.duckdb",
  destfile = "flights.duckdb",
  mode = "wb")

# Connecting to it
conn <- dbConnect(duckdb(), dbdir = "flights.duckdb", read_only = TRUE)

# Verifying connection
dbListTables(conn)
```

#### a. List the first 10 rows of the flights table.

```{r}
dbGetQuery(conn, "
  SELECT *
  FROM flights
  LIMIT 10
")
```

#### **b.**

Find the **number of flights**

i.  in the dataset

ii. for each carrier

```{r}
# part i
dbGetQuery(conn, "
  SELECT COUNT(*) AS total_flights
  FROM flights
")

# the total number of flights in the dataset is 336,776.
```

```{r}
# part ii
dbGetQuery(conn, "
  SELECT carrier, COUNT(*) AS num_flights
  FROM flights
  GROUP BY carrier
  ORDER BY num_flights DESC
")

# the total number of flights for each carrier can be seen in the table below:
```

#### **c.**

Find the **average departure delay** (`dep_delay`)

```{r}
dbGetQuery(conn, "
           SELECT AVG(dep_delay) AS avg_dep_delay
           FROM flights
           ")

# The average departure delay is 12.64 minutes (rounded to the nearest hundredth).
```

#### **d.**

Find the **top 5 destinations** (`dest`) With the most flights.

-   Your output should include:

    -   destination

    -   number of flights Sort from the **largest to smallest**.

```{r}
dbGetQuery(conn, "
           SELECT dest, COUNT(*) AS num_flights
           FROM flights
           GROUP BY dest
           ORDER BY num_flights DESC
           LIMIT 5
           ")

# the top 5 destinations with their corresponding number of flights can be seen below, sorted from largest to smallest:
```

#### **e.**

Create a table showing:

-   `carrier`

-   average arrival delay (`arr_delay`)

Sort the results from **largest to smallest delay**

```{r}

```

> You must close the connection when you are done. Use `dbDisconnect()`

```{r}

```

### **Part II: Answer the following question by running SQL using DBeaver with DuckDB.**

> Please go over the note **dbeaver_sql.html** to learn

#### **a.**

List the first **10 rows** of the `flights` table.

```         
dbGetQuery(conn, "YOUR SQL HERE")
```

```{r}

```

#### **b.**

Count the Number of Flights for Each Carrier

Find the **number of flights** for each carrier.

-   Your output should include:

    -   carrier

    -   number of flights Sort from the **largest to smallest**.

```{r}

```

#### **c.**

Find the **average departure delayby month**

-   Your output should include:

    -   month

    -   average departure delay

```{r}

```

#### **d.**

Find the **top 10 routes** (`dest`) by:

-   `origin`

-   `dest`

ordered by **number of flights**.

```{r}

```

#### **e.**

Export the result of part b query to a **CSV file** using the **Export Data** feature in DBeaver.

Submit the exported CSV file along with your homework.

-   **Exporting Data to CSV in DBeaver**

```{r}

```

**Quick Steps:**

1.  **Right-click** Table or Result Grid -\> **Export Data**.

2.  Select **CSV** -\> Click **Next**.

3.  Confirm **Settings** (Delimiters, Encoding) -\> Click **Next**.

4.  Choose **Output Directory** -\> Click **Finish**.

## **Remember to Close the connection When You Are Done**

-   **How to Disconnect and Reconnect a Database in DBeaver**

1.  On the **top menu**, click **Database**.

2.  From the drop-down menu, select **Disconnect**.

3.  You should see a message on the **bottom right corner**:

    ```         
    Not connected to database
    ```

4.  To reconnect, go to **Database → Connect** and select your connection.

> This is the easiest way to manage connections without closing DBeaver or losing work.

```{r}

```

## **Submission**

-   Save your Quarto file after completing each problem, render it, then stage and commit the changes.

-   Include well-commented code and clear explanations.

-   Make appropriate **commits**.

-   Submit the rendered file to **Canvas** and push it to your **GitHub repository**.

**Good luck!**
